### 2020-08-10 thesis/thesis.tex
fix linking statement
gen disc

\SetLinkName{Linking Statement}
\SetLinkText{In Chapter I, we integrated a land use model and a climate change model with a habitat suitability and connectivity model. We then explored outcomes under different scenarios for land use and climate change. We showed how different species might react to changes in land use and climate by analyzing the resulting flow surfaces. We provided  a reproducible yet perfectible methodology that takes into account important drivers like land use and climate change. We applied these methods to the case of the Montérégie in Québec.. We demonstrate the power of simple scenarios to identify the boundaries of the prediction space for potential functional connectivity. However, we recognize the need for scenarios that take into account the needs and perceptions of stakeholders of the region. In Chapter 2, we take a first step toward deriving the basis for such community-driven scenarios by presenting the results of a community workshop organized between multiple stakeholders in Montérégie. We use those results to derive simple land use change scenarios and use those scenarios to constrain the integrated model from Chapter 1. These scenarios can serve as a basis with which to iterate a community-driven process of conservation scenario building. We derive the same analysis for the species described in Chapter 1 and compare and contrast the results between scenarios. This integrated approach has the potential to support ongoing connectivity conservation planning in the south of Québe}
\Link

@@

% General Discussion
% \include{discussion}
\SetLinkName{{General Discussion \& Conclusion}}%
% In case of a manuscript-based thesis the comprehensive discussion should encompass all of the chapters of the thesis and should not be a repetition of the individual chapters.
\SetLinkText{TEST ---- A classic spatially defined land use change model is what can be called a “state change model”, where the landscape is divided in a grid and each grid element (pixel) is assigned a state that changes in discrete time \citep{frazer_model_1984}. The difficulty of land use change modelling resides in setting the transition probabilities that define the probability of a pixel transitioning to another state (or to remain as is), given its current state and given a set of conditions both intrinsic and extrinsic to that pixel. --- TEST}%
\Link%

>>>>

\SetLinkName{Linking Statement}
\SetLinkText{In Chapter I, we integrated a land use model and a climate change model with a habitat suitability and connectivity model. We then explored outcomes under different scenarios for land use and climate change. We showed how different species might react to changes in land use and climate by analyzing the resulting flow surfaces. We provided  a reproducible yet perfectible methodology that takes into account important drivers like land use and climate change. We applied these methods to the case of the Montérégie in Québec.. We demonstrate the power of simple scenarios to identify the boundaries of the prediction space for potential functional connectivity. However, we recognize the need for scenarios that take into account the needs and perceptions of stakeholders of the region. In Chapter 2, we take a first step toward deriving the basis for such community-driven scenarios by presenting the results of a community workshop organized between multiple stakeholders in Montérégie. We use those results to derive simple land use change scenarios and use those scenarios to constrain the integrated model from Chapter 1. These scenarios can serve as a basis with which to iterate a community-driven process of conservation scenario building. We derive the same analysis for the species described in Chapter 1 and compare and contrast the results between scenarios. This integrated approach has the potential to support ongoing connectivity conservation planning in the south of Québec.}
\Link

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%        General Discussion                     %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Clearly state the rationale and objectives of the research.
\newcommand*{\SetDiscName}[1]{\renewcommand*{\ETDDiscName}{#1}}%
\newcommand*{\ETDDiscName}{DiscMENTS}%

\newcommand{\SetDiscText}[1]{\renewcommand*{\ETDDiscText}{#1}}%
\newcommand*{\ETDDiscText}{Disc text goes here!}%

\newcommand{\Disc}{%
    \begin{simpleenv}{}{}{}{}%
    \pagestyle{plain}%
  \setboolean{SetDSpace}{false}%
    \GoSingle%
    \begin{QZ@Cent}%
          \bfseries{\ETDDiscName}
      \end{QZ@Cent}%
    \vspace*{0.5in}
      \par%
      \GoDouble%
      \setlength{\parskip}{-2em}
      %\chapter*{\ETDDiscName}%
      \addcontentsline{toc}{extrachapter}{\ETDDiscName}%
      \ETDDiscText%
    \end{simpleenv}%
    \setboolean{SetDSpace}{true}}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Discussion
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SetDiscName{General Discussion \& Conclusion}%

\SetDiscText{
The two chapters of this thesis each provide parts of a strategy for addressing the two limitations to Connectivity Conservation Planning methods identified in the general introduction: the lack of methodological integration between connectivity, land use change and climate change models ; and the need for stakeholder inputs in the planning process. \\

In chapter 1, we integrated circuit-theory based connectivity models with a simple land use and climate change model based on habitat suitability for 5 umbrella species. We also showed patterns of land use change in our study region for the 1990-2010 period, and demonstrated that our land use change model is a generalization of these patterns across the landscape. We analysed changes in connectivity under simple land use and climate change scenarios, and showed that different species have different responses, depending on the forest type they depend on, and on whether this forest type is under threat by land use change, climate change or both. The american marten showed the most complex response because it is dependent on coniferous forest type which is the most at risk under both drivers. We also used a pattern-detection method (SURF) to identify pinch-points and salient connectivity flow features. Although the number of features detected is species-specific, we showed a mostly consistent pattern of change over time: reforestation seems to simplify the flow. More research to understand the links between flow complexity and SURF measures is needed. \\

Then, in chapter 2, we capitalized on chapter 1’s progress by including the results of a stakeholder workshop during which we mapped consensual areas of importance for connectivity in the region. During this participatory process, we also identified key assets (such as the drive and momentum of conservation NGOs in the region) and obstacles (such as the predominance of agricultural matrix) to connectivity conservation in the region. We derived simple conservation scenarios based on priorities identified during the participatory mapping exercise. We simulated connectivity changes for our 5 species under these scenarios and put in evidence the complexity of restoring connectivity to cater to species with different needs while accounting for multiple stakeholder interests. \\

Some general limitations that concern both chapters have to be addressed. We argued that using markov-chain monte carlo (MCMC) based modelling for land use change allowed us to take into account the uncertainty of land use change simulations. The principle is to run the model multiple times (iterations), resulting in slightly different simulations every time. This allowed us to see that our model was quite constrained and that over 10 iterations we produced narrow confidence intervals when analysing the resulting connectivity, in both chapters. However, there remains a large amount of uncertainty unaccounted for in the model. STSIM required a large number of parameters (for example, neighboring parameters, see \ref{tab:neigh_rules}) that were chosen and fixed to a certain value. To correctly account for uncertainty in those parameters, we could decide on a distribution for those parameters and draw from this distribution at each iteration. This is possible in STSIM, which allows the user to specify distributions. However this brings its own set of drawbacks. First, in order to get a representative sample from this distribution, more iterations would need to be computed, which would increase computing time. Second, the shape of the distribution for parameters like neighboring rules is not easy to determine a priori. This last point could be alleviated by bootstrapping the distribution from data. \\

Our approach to modeling functional connectivity is also limited. In both chapters, we use circuit theory to model potential functional connectivity (PFC). A limitation of PFC modelling is that what the model produces is an informed prediction about potential movement, which is still in need of statistical validation. Therefore, the quality of the data used for parameterization is extremely important for the accuracy of the results. In addition, circuit theory has been criticized as a PFC approach. \cite{fletcher_spatial_2018} mention a set of issues: first, animals might not move in a way that can realistically be modelled as random walks. It is computationally more expensive than least-cost path analysis - although the recent Julia version of Circuitscape is faster and more efficient than previous iterations of the software. They also argue that results can be difficult to interpret across large landscapes, given the fact that flow tends to be predicted as being diffuse. Finally, they mention an issue that “has been neglected in the ecological literature but has been repeatedly acknowledged in the network literature.” According to \cite{luxburg_hitting_2014}, on large graphs, the resistance distance  between patches converges on the local properties at a cell level, making the use of random walks less helpful. \\

Finally, both chapters point out limitations in the way we defined land use change and conservation scenarios. Although the goal was to derive simple and extreme cases to show the width of variations in species response, this method will remain unsatisfactory for stakeholders, until it is inscribed in an interactive process of community-driven scenario building we pointed to at the end of Chapter 2. Both chapters outline the design principles for a tool that could be used to facilitate this iterative process, by providing information on the consequence of “what if” scenarios on potential functional connectivity. It is important to ask what framework would best allow us to operationalize this tool and allow researchers to provide guidance to stakeholders during a decision making process. In the following, we discuss the importance of the Spatialized SES (SpaSES, \cite{williamson_spatially_2018}), and Ecosystem Services Bright Spots \citep{frei_bright_2018} frameworks to guide the design of connectivity conservation planning tools. \\

Having recognized that conservation tools often fail to consider social factors affecting their implementation, \cite{williamson_spatially_2018} developed the SpaSES framework as a recontextualization of Ostrom’s framework applied to conservation tools. In the original SES framework, SES variables are numerous, and acquiring data on them for decision making at large spatial scales is often impossible. Therefore, \citeauthor{williamson_spatially_2018} bundled them in three categories, forming a three dimensional space across which probability of conservation action can be modelled. First, Ecological Value (EV) describes how important an habitat is for conservation. Then, Social Willingness (SW) and Institutional Capacity (IC) account for social factors. Social willingness variables essentially take stock of the level of social acceptability to conservation, based on theories of individual and collective behavior. Institutional Capacity variables measure governance systems, institutional fit, and factors that influence the capacity of institutions of different types (government, action groups) to reach conservation goals. \citeauthor{williamson_spatially_2018} models probability of connectivity as a function of those three latent variables (EV, SW, IC), for which they suggest a number of proxies that will vary depending on study region and conservation objectives. \\

For our study region, an appropriate measure of EV could be the results of a prioritization of habitat patches based on habitat suitability and connectivity as different scales under land use and climate change, such as in \cite{albert_applying_2017}. But \citeauthor{williamson_spatially_2018} notes that this type of prioritization effects “generally assume that stakeholders are interested in an optimal solution to resource management concerns and pursue those solutions rationally, ” an assumption that is not confirmed by SES research. This justifies the need for taking SW and IC into account. Participatory methods such as the one we employed could be an appropriate way of developing measures for these two axes, as stakeholders most likely know what variables would be better representative of each axis. An iterative process of scenario building could be initiated by efforts to define (1) multiple scenarios based on stakeholder’s perceptions of planning priorities, (2) a measure of SW for conservation, partially based on alignment with the different scenarios developed and (3) a measure of stakehodelrs’s confidence in IC in the region to conduct different conservation plans, and an idea of what the needs are in IC for such plans to succeed. This framework would lead to better estimates of where conservation is both critical and implementable in the region. \\

Workshops might also be a good tool to build up SW and IC themselves and increase the probability of conservation action, which will in a large part depend on addressing the obstacles identified in our workshop. According to our results, the most obvious of these obstacles in our region is the predominance of agricultural land use. The role of the agricultural matrix in fragmented landscapes has been under scrutiny at least since the theory of island biogeography and the SLOSS debate in ecology, which was for a time renewed in the land sparing versus land sharing debate. The concept at the base of these dichotomies, is that smaller fragments experience higher extinction, and that connectivity can help by restoring immigration between fragments. \cite{perfecto_natures_2019} points out that focusing on connectivity and immigration rates in fragmented landscapes actually “means focusing on the agroecosystems, which in turn means examining what is happening socio-politically in the agricultural matrix” \citep{perfecto_natures_2019}. Both the “single large/several small” and “sharing/sparing” oppositions are now seen as false dichotomies, rendered obsolete by the SES literature. \cite{bennett_changing_2017} agrees and argues that those dichotomies have tended to look at agrosystems too simply, mostly within the angle of food production. They propose we go beyond those dichotomies by acknowledging the multiple facets of agroecosystems and considering the complexity of the agro-ecological matrix. \\

One way to embrace the complexity of the matrix is to carefully look at the relationship between the structure of a landscape and its impact on ecosystem functions. Research has shown that landscape complexity can increase biodiversity. For example, crop heterogeneity is linked to higher biodiversity \citep{sirami_increasing_2019}. Agricultural matrices with smaller average field sizes tend to be more biodiverse \citep{fahrig_farmlands_2015}. These phenomena have been attributed to an increase in connectivity due to the presence of more semi-natural habitats. These contributions are important and should be considered for land use planning. They are appealing for the design of connectivity conservation tools, because they point out clear management scenarios that could be tested through models, and can make use of the increasing availability of high resolution remotely sensed data. \\

But, as we quoted from \cite{williamson_spatially_2018} above, this view tends to assume that stakeholders want to pursue rational and optimal design in their approach to conservation, and is once again a single-faceted approach to the problem, focusing on the trade-off between food production and support biodiversity. An alternative way is through the lens of ecosystem services  \cite{bennett_changing_2017}. Ecosystem services refer to the diversity of benefits that ecosystems provide to, and the many ways in which they support, human systems. Specifically, the concept of “bright spots” has potential for guiding the discussion beyond dichotomies and single faceted approaches, by focusing on multifunctionality in agroecosystems. In a landscape, “bright spots” are areas exceeding expectations for goals, such as agro-ecological multifunctionality, biodiversity, or connectivity \citep{frei_bright_2018}. They show promise as a tool  to identify levers of change, and guide stakeholder discussion. Indeed, an approach that focuses on identifying bright spots for the provision of those ecosystem services can be valuable to an interactive process of stakeholder engagement. They offer starting points to work with to design scenarios. They are concrete examples to reason around what bundles of ecosystem services are desirable (elements of SW) and which types of governance is needed to work toward them (elements of IC). In our study region, this framework is already being tested: \cite{frei_bright_2018} identified bright, dark and average spots within Monteregie for both landscape multifunctionality and avian biodiversity. They showed important spatial properties of bright spots, for instance that they were “neither spatially congruent nor in conflict”. \\

\citeauthor{frei_bright_2018} also pointed out a range of challenges for the identification of bright spots, the development of multicriteria measures being one of them. Connectivity measures are already being developed with  multifunctionality in mind. For instance, \cite{rayfield_priorisation_2018}, assessed whether certain habitat networks in the Monteregie would be more resilient to future change using several multi-scale metrics of connectivity. Another example is \cite{fletcher_towards_2019} who developed an unifying model for connectivity that takes into account multiple types of motability. Connectivity is therefore a field that fits in the multicriteria framework of bright spots, as it itself moves in this direction.  Therefore bright spots show promise for being integrated in more complex connectivity conservation planning tools. Finally, bright spots are intuitive, being based on statistical deviation from a given distribution and are therefore always relative to the context in which they are implemented. \\

\vspace{3em}

\subsection*{\textit{Conclusion \\ \vspace{2em}}}

The integration of models of multiple drivers of changes in ecological connectivity with stakeholder inputs into connectivity models will contribute to the design of better connectivity conservation planning tools. In this thesis, we contributed some design elements for those tools, such as a philosophy to account for uncertainty and to confront models with the priorities of local stakeholders. Future research will need to engage with an iterative process of stakeholder engagement, possibly a scenario building exercise. This process will be all the more successful if it is made within a coherent framework that addresses the social-ecological complexity of fragmented landscapes, such as SpaSES and Bright spots. These frameworks will guide discussion and facilitate visualizing positive alternative scenarios to business as usual.
}%
\Disc%
